\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[margin=1.25cm,top=1.35cm,bottom=1.25cm]{geometry}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{titlesec}

\definecolor{accent}{HTML}{0f3460}
\definecolor{dark}{HTML}{1a1a2e}
\definecolor{softgreen}{HTML}{e8f5e9}
\definecolor{midgray}{HTML}{666666}

\hypersetup{colorlinks=true,linkcolor=accent,urlcolor=accent}
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}

\titleformat{\section}{\normalsize\bfseries\color{accent}}{}{0em}{}
\titlespacing{\section}{0pt}{4pt}{2pt}

\begin{document}

\begin{center}
{\LARGE\bfseries\color{dark} XC Gradient}\\[0.1em]
{\small\color{midgray} Executive One Pager $\cdot$ April 2026 $\cdot$ Confidential}\\[0.35em]
\colorbox{accent}{\color{white}\small\bfseries\strut\quad We reduce industrial fault diagnosis from hours to minutes, with private on-premises AI \quad}
\end{center}

\vspace{0.25em}
{\color{accent}\hrule height 0.8pt}
\vspace{0.35em}

\begin{multicols}{2}

\section{Problem}
Manufacturing SMEs hold critical know-how across PDFs, internal procedures, and senior technicians' memory. When a machine fails, the bottleneck is not the repair itself: \textbf{it is diagnosis}.

\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item High MTTR due to manual information retrieval.
\item Dependency on key experts approaching retirement.
\item Sensitive operational data that cannot be pushed to public cloud AI.
\end{itemize}

\section{Solution}
XC Gradient deploys a private AI layer inside each plant:
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item \textbf{Diagnostic copilot}: actionable fixes in seconds from manuals, incidents, and logs.
\item \textbf{Real-time OEE}: availability, performance, and quality per machine on one operational dashboard.
\item \textbf{Automatic structuring}: unstructured documentation becomes a searchable knowledge base.
\end{itemize}

\begin{center}
\colorbox{softgreen}{\parbox{0.94\linewidth}{\centering\small\bfseries
We sell recovered productive time.\\
The value is fewer stoppages, not more dashboards.
}}
\end{center}

\section{Why now}
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item \textbf{NIS2 (2024--2025)} increases pressure for industrial data sovereignty.
\item \textbf{Senior workforce retirement wave} across EU manufacturing.
\item \textbf{SME digitization momentum} now supports vertical software adoption.
\end{itemize}

\section{Early traction}
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item \textbf{Decfa} (CNC, $\sim$30 employees): PoC in execution.
\item \textbf{Paver} (Granollers): PoC in demo phase.
\item Active pipeline with additional industrial SMEs in Catalonia.
\end{itemize}

\columnbreak

\section{Market (YC: big pain + big market)}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabularx}{0.96\linewidth}{lX}
\toprule
\textbf{TAM} & $\sim$2M manufacturing SMEs in the EU \\
\textbf{SAM} & $\sim$180k in Spain (20--200 employees) \\
\textbf{SOM} & $\sim$500 reachable companies (Catalonia, years 1--2) \\
\bottomrule
\end{tabularx}
\normalsize

\section{Business model}
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item B2B subscription (per machine or tiered SaaS, under PoC validation).
\item \textbf{Land-and-expand}: start with a critical workflow, expand plant-wide.
\item Target SaaS gross margin: \textbf{$>$70\%}.
\end{itemize}

\section{Defensible advantage}
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item \textbf{Structural privacy}: on-prem inference, zero cloud dependency.
\item \textbf{Customer ontology accumulation}: each deployment increases switching costs.
\item \textbf{Proprietary IP}: Markov+Choquet reranking and multi-tenant S-QDoRA fine-tuning.
\end{itemize}

\section{Team}
\begin{itemize}[leftmargin=1em,itemsep=1pt,topsep=2pt]
\item \textbf{Oriol Farr\'es} (CEO/CFO): core AI architecture and strategy.
\item \textbf{Arnau Noguer} (CTO/AI): RAG, ontologies, and document ingestion.
\item \textbf{Adam Sarrate} (COO/CISO): infrastructure, security, and deployment.
\end{itemize}
UPC-FIB engineering founders, full-time.

\section{Funding path and next milestone}
Bootstrapped with grants (Santander X, Impuls UPC). \textbf{Seed round after 100+ paying customers}, backed by real metrics (MTTR, retention, ARR).

\end{multicols}

\vspace{0.2em}
{\color{accent}\hrule height 0.8pt}
\vspace{0.15em}
\begin{center}
\small\textcolor{midgray}{XC Gradient $\cdot$ Founder-led contact $\cdot$ No public website yet $\cdot$ Confidential}
\end{center}

\end{document}
