\chapter{Clustering Utility Study}
\label{chap:clustering-introduction}

\section{Purpose of the Study}
\label{sec:clustering-purpose}

Part I evaluates synthetic-data utility in the unsupervised-learning setting. Within the general thesis framework, clustering is used as a test of \textbf{structural utility}: whether a synthetic dataset preserves the geometry, separability, density, and dispersion required for an algorithm to recover meaningful groups.

The study focuses on synthetic data generated through the Classification and Regression Trees (CART) method implemented in the \texttt{synthpop} package. CART is flexible, but its axis-aligned tree partitions can turn smooth multivariate structure into block-like regions. This potential artifact is directly relevant to clustering because clustering algorithms depend on distances, shapes, and local connectivity.

\section{Clustering as a Structural Utility Test}
\label{sec:clustering-structural-utility}

Unlike supervised methods, clustering algorithms do not have access to ground-truth labels during training. Their objective is to discover inherent structures by grouping similar observations together based on geometric or statistical properties.

Clustering is fundamental for exploratory data analysis, customer segmentation, and anomaly detection. However, because these algorithms rely heavily on the geometric structure of the data (distances, densities, and distributions), they are highly sensitive to data quality and distribution. Common algorithms include:
\begin{itemize}
  \item \textbf{K-Means:} Partitioning data into $k$ groups by minimizing variance within clusters.
  \item \textbf{Hierarchical Clustering:} Building a tree of clusters (dendrogram) based on distance connectivity.
\end{itemize}

This part of the thesis extends the work of Chen Xinnuo \cite{chen2025} by systematically comparing how well K-Means and Agglomerative Hierarchical Clustering perform on CART-generated synthetic data.

We aim to answer three specific research questions:
\begin{enumerate}
  \item \textbf{Detection Capability:} Can clustering algorithms correctly identify the optimal number of clusters ($k$) in synthetic data as well as they do in real data?
  \item \textbf{Algorithm Robustness:} Which algorithm is more resilient to the distortions introduced by the synthetic generation process? We hypothesise that Hierarchical Clustering may be more robust than K-Means in non-normal distributions.
  \item \textbf{Quantification of Degradation:} To what extent does the CART generation method degrade the geometric properties (separability, density) of the clusters?
\end{enumerate}

To answer these questions, we have developed a reproducible pipeline that generates 1,800+ datasets varying in sample size ($N$), cluster separation, noise levels, and probability distributions (Normal vs. Gamma), allowing for a rigorous statistical comparison.

\section{Structure of Part I}
\label{sec:clustering-structure}
This part is organized as follows:
\begin{itemize}
  \item \textbf{Chapter 3: Methods} details the mathematical formulations of K-Means and Hierarchical Clustering and describes the synthetic data generation via CART. It also defines the structural fidelity metrics and the simulation pipeline used to evaluate algorithm robustness.
  \item \textbf{Chapter 4: Results} presents the empirical analysis, including PCA and t-SNE visualizations alongside cluster detection success rates. It further quantifies geometric degradation using the Gini coefficient, Mean Centroid Distance, and Mean Variance Difference metrics.
  \item \textbf{Chapter 5: Conclusions} synthesizes the clustering findings, highlighting the superior robustness of Hierarchical Clustering for skewed synthetic data and identifying the practical implications for structural utility evaluation.
\end{itemize}
